---
title: "Coal Exploration"
author: "Peter Boshe"
date: "2022-06-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r}
# load libraries
library(tidyverse)
library(ggridges)
library(GGally)
library(janitor)

# parameters
  #input
file_raw <- here::here("data/clean_coal.rds")

# set theme
theme_set(theme_bw())

```



```{r}

# load the dataset and explore
clean_coal <- read_rds(file_raw)
clean_coal |> 
  ggpairs(columns = c(2,5,6,7,8,9),
          mapping = ggplot2::aes(color = commodity),
          diag = list(alpha = 0.5))
```



```{r}
clean_coal_sum <- clean_coal |> 
  pivot_wider(names_from = trade,
              values_from = value_usd) |>
  replace_na(list(imp = 0,
                  exp = 0)) |>
  group_by(year) |> 
  summarise(total_exp = sum(exp),
            total_imp = sum(imp),
            count = n())

# to be joined to the original dataset

clean_coal_sum |> 
  pivot_longer(cols = starts_with("total_"),
               names_to = "trade",
               values_to = "value",
               names_prefix = "total_") |> 
  ggplot(aes(year, value, fill = trade)) + 
  geom_col(position = "dodge",
           stat = "identity",
           alpha = 0.8) + 
  scale_y_log10(labels = scales::dollar)


```

## import and export before 2012


```{r}
clean_coal_sum |> 
  filter(year< 2012) |> 
  summarise(total_imp = sum(total_imp),
            total_exp = sum(total_exp))
```
## total number import and export

```{r}
clean_coal_sum |>
  summarise(total_imp = sum(total_imp),
            total_exp = sum(total_exp))
```



## trade value import and export

```{r}

clean_coal |> 
  filter(value_usd > 1 ) |> 
  ggplot(aes(value_usd, fill= trade, color = trade)) +
  geom_histogram(position = "identity",
                 alpha = 0.5,
                 binwidth = 1) +
  scale_x_log10(labels = scales::dollar)
```


## commodity, years, trade value


```{r}


clean_coal |> 
  ggplot(aes(value_usd, year, fill = trade, group = year)) +
  geom_density_ridges(scales = 1,
                      alpha = 0.7) +
  scale_x_log10(labels = scales::dollar) +
  facet_wrap( ~ trade)# TODO to animate graphs above
```


## trade, commodity

```{r}
clean_coal |> 
  tabyl(trade,commodity)
# FIX cleaning script must have an issue, consider building a data validator
# maybe we should have pivoted longer before ommiting NAs
```
 
 - we have a problem here, fixed
 
 
## use code as startpoint to separate the trading instances
 
```{r}
clean_coal |> 
  tabyl(trade,partner_iso)

#to simply have a dataframe of export countries and one for import countries
# the 
```
 

## coal rent, years

```{r}
clean_coal |> 
  ggplot(aes(year, coal_rent_pct_of_gdp)) + 
  geom_line()
```


## coal rent, trade



## natural capital, years

```{r}
clean_coal |> 
  ggplot(aes(year, natural_capital)) + 
  geom_line()
```


## coal rent, natural capital


```{r}
clean_coal |> 
  ggplot(aes(coal_rent_pct_of_gdp, natural_capital, color = as.factor(year))) + 
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm",
              color = "darkgrey",
              lty = 2)

```

## network analysis/connection map

### method 1

```{r}
library(countrycode)
library(rnaturalearth)
library(raster)

# load in map object
map <- ne_countries()

# add the trade value and import export categories to the map obj
map$trade <- as_vector(clean_coal[match(map$iso_a3,str_to_upper(clean_coal$partner_iso)), "trade"])
map$value_usd <- as_vector(clean_coal[match(map$iso_a3,str_to_upper(clean_coal$partner_iso)), "value_usd"])

plot(map)


```


### method 2

```{r}

library(maps)
library(geosphere)
# import iso2c, country name, longitude and latitude data

countries_clean_coal <- clean_coal |> 
  mutate(partner = map_chr(partner_iso, ~ countrycode(., origin = "iso3c", destination = "country.name")),
        reporter = map_chr(reporter_iso, ~ countrycode(., origin = "iso3c", destination = "country.name")))

partner_countries <- countries_clean_coal |> 
  distinct(partner_iso, partner)

reporter_countries <- countries_clean_coal |> 
  distinct(reporter_iso,reporter)

# import longitude and latitude dataframe
countries_metadata <- read_csv("https://gist.githubusercontent.com/tadast/8827699/raw/f5cac3d42d16b78348610fc4ec301e9234f82821/countries_codes_and_coordinates.csv")

#clean the imported dataframe
countries_metadata <- countries_metadata |> 
  mutate_if(is.character, str_to_lower) |> 
  clean_names()  |> 
  dplyr::select(-numeric_code) |> 
  rename(latitude = latitude_average,
         longitude = longitude_average)

#create data frames for the partner and reporter countries
partner_coord <- partner_countries |> 
  left_join(countries_metadata, by = c("partner_iso" = "alpha_3_code"))

reporter_coord <- reporter_countries |> 
  left_join(countries_metadata, by = c("reporter_iso" = "alpha_3_code"))

# create a base map
map("world", 
    fill=T, 
    col="grey8", 
    bg="grey15")

# overlay the traders

points(partner_coord$longitude, 
       partner_coord$latitude,
       pch = 3,
       cex=0.1,
       col = "red")

# create the connections
for (i in (1:dim(partner_coord)[1])) { 
inter <- gcIntermediate(c(reporter_coord$longitude[1], reporter_coord$latitude[1]), c(partner_coord$longitude[i], partner_coord$latitude[i]), n=200)
lines(inter, lwd=0.5, col="yellow")    
}


```

## stickman figures

```{r}
library(sysfonts)
library(ggplot2)
library(xkcd)
font.add("xkcd", regular = "xkcd.ttf")


ratioxy <- diff(xrange)/diff(yrange)
mapping <- aes(x, y,
scale,
ratioxy,
angleofspine ,
anglerighthumerus,
anglelefthumerus,
anglerightradius,
angleleftradius,
anglerightleg,
angleleftleg,
angleofneck)

dataman <- data.frame(x= 15, y= 4,
scale = 0.4,
ratioxy = ratioxy,
angleofspine = -pi/2 ,
anglerighthumerus =  -pi/6,
anglelefthumerus = pi + pi/6,
anglerightradius = -pi/3,
angleleftradius = -pi/3,
anglerightleg = 3*pi/2 - pi / 12,
angleleftleg = 3*pi/2 + pi / 12 ,
angleofneck = runif(1, 3*pi/2-pi/10, 3*pi/2+pi/10))

g <- ggplot() + 
  # xkcdaxis(xrange,yrange) +
  xkcdman(mapping,linetype = 3, dataman) + 
  theme_void() +
  annotate("text", x = 17, y = 5, label = "Some text", size = 7, family = "xkcd") +
  theme(line = element_blank(),
        text = element_text('xkcd')) +
  annotate("rect", xmin = 14.5, xmax = 19.5, ymin = 4.5, ymax = 5.3,
  alpha = .2) +
  annotate("segment",x = 16.2, xend = 16.6 ,
             y = 4.2, yend = 4.4, curvature = 1) 
g
```



















