---
title: "Coal Story, Bro..."
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme:
      version: 4
      bootswatch: sketchy 
---

```{=html}
<style>
.sbframe-component {
  background-color: #D3D3D3;
}
</style>
```
```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(bslib)
library(ggridges)
library(janitor)
library(maps)
library(geosphere)
library(countrycode)
library(thematic)
library(sysfonts)
library(ggplot2)
library(xkcd)
library(patchwork)
library(ggtext)

# parameters
  #input
file_raw <- here::here("data/clean_coal.rds")

# set theme
theme_set(theme_bw() + 
            theme(panel.background = element_rect(fill = "white")))

# add fonts
font_add("xkcd", regular = "xkcd.ttf")

# load the dataset
clean_coal <- read_rds(file_raw)

```

### Introduction

<center>

<h4>**Tanzania Coal Trade**: the disparity between [**imports**]{style="color:yellow"} and <span style='color:green'>**exports**<span>.</h4>

</center>

```{r, fig.height=6, fig.width=19}

thematic::thematic_off()

countries_clean_coal <- clean_coal |> 
  mutate(partner = map_chr(partner_iso, ~ countrycode(., origin = "iso3c", destination = "country.name")),
        reporter = map_chr(reporter_iso, ~ countrycode(., origin = "iso3c", destination = "country.name")))

partner_countries <- countries_clean_coal |> 
  distinct(partner_iso, partner)

reporter_countries <- countries_clean_coal |> 
  distinct(reporter_iso,reporter)
# TODO need to import and save this dataset
# import longitude and latitude dataframe
countries_metadata <- read_csv("https://gist.githubusercontent.com/tadast/8827699/raw/f5cac3d42d16b78348610fc4ec301e9234f82821/countries_codes_and_coordinates.csv")


#clean the imported dataframe
countries_metadata <- countries_metadata |> 
  mutate_if(is.character, str_to_lower) |> 
  clean_names()  |> 
  dplyr::select(-numeric_code) |> 
  rename(latitude = latitude_average,
         longitude = longitude_average)

#create data frames for the partner and reporter countries
partner_coord <- partner_countries |> 
  left_join(countries_metadata, by = c("partner_iso" = "alpha_3_code")) |> 
  distinct(latitude,longitude, .keep_all = TRUE) 

#to simply have a dataframe of export countries and one for import countries
partner_imp <- clean_coal |> filter(trade == "imp") |> distinct(partner_iso)
partner_exp <- clean_coal |> filter(trade == "exp") |> distinct(partner_iso)

#separate the data by trade type
partner_coord_imp <- partner_imp |> left_join(partner_coord, by = "partner_iso") 
partner_coord_exp <- partner_exp |> left_join(partner_coord, by = "partner_iso") 

reporter_coord <- reporter_countries |> 
  left_join(countries_metadata, by = c("reporter_iso" = "alpha_3_code"))

# create a base map
map("world", 
    fill=T, 
    col="grey8", 
    bg="grey15")

# overlay the traders

points(partner_coord$longitude, 
       partner_coord$latitude,
       pch = 3,
       cex=0.1,
       col = "red")

# create the connections for imports
for (i in (1:dim(partner_coord_imp)[1])) { 
inter <- gcIntermediate(c(reporter_coord$longitude[1], reporter_coord$latitude[1]), c(partner_coord_imp$longitude[i], partner_coord_imp$latitude[i]), n=200)
lines(inter, lwd=1, col="yellow")    
}

# create the connections for export
for (i in (1:dim(partner_coord_exp)[1])) { 
inter <- gcIntermediate(c(reporter_coord$longitude[1], reporter_coord$latitude[1]), c(partner_coord_exp$longitude[i], partner_coord_exp$latitude[i]), n=200)
lines(inter, lwd=1, col="green")    
}


```

### Trade value over the years

```{r}
# set document theme

thematic::thematic_off()


xrange <- c(0,20) # might need to change range
yrange <- c(0,30) # might need to change range
ratioxy <- diff(xrange)/diff(yrange)
mapping <- aes(x, y,
scale,
ratioxy,
angleofspine ,
anglerighthumerus,
anglelefthumerus,
anglerightradius,
angleleftradius,
anglerightleg,
angleleftleg,
angleofneck)

dataman <- data.frame(x= 0, y= 2,
scale = 2,
ratioxy = ratioxy,
angleofspine = -pi/2 ,
anglerighthumerus =  -pi/16,
anglelefthumerus = pi + pi/6,
anglerightradius = pi/2 - pi/4,
angleleftradius = pi - pi/4,
anglerightleg = 3*pi/2 - pi / 12,
angleleftleg = 3*pi/2 + pi / 12 ,
angleofneck = runif(1, 3*pi/2-pi/10, 3*pi/2+pi/10))

 g <- ggplot() + 
  # xkcdaxis(xrange,yrange) +
  xkcdman(mapping,linetype = 3, dataman) + 
  theme_void() +
  # this will allow the text outside of the plot panel
  coord_cartesian(clip = 'off',
                  expand = FALSE) +
  annotate("richtext", x = 3, y = 5.4, ymax = 10,xmin = -4,xmax = 10,  label = "The total value<br>of <span style='color:green'>exports</span> had never <br>exceeded the $100,000 <br>mark prior to 2013, <br>while <span style='color:yellow'>imports</span> have always <br>been up there..", family = "xkcd",
           fill = "lightgrey") +
  theme(line = element_blank(),
        text = element_text('xkcd')) +
  # annotate("rect", xmin = 2, xmax = 5, ymin = 4.5, ymax = 10,
  # alpha = .2) +
  annotate("segment",x = 0.9, xend = 1.5 ,
             y = 2.5, yend = 3.2) 



clean_coal_sum <- clean_coal |> 
  pivot_wider(names_from = trade,
              values_from = value_usd) |>
  replace_na(list(imp = 0,
                  exp = 0)) |>
  group_by(year) |> 
  summarise(total_exp = sum(exp),
            total_imp = sum(imp),
            count = n())

# to be joined to the original dataset

h <- clean_coal_sum |> 
  pivot_longer(cols = starts_with("total_"),
               names_to = "trade",
               values_to = "value",
               names_prefix = "total_") |> 
  ggplot(aes(year, value, fill = trade)) + 
  geom_col(position = "dodge",
           stat = "identity",
           alpha = 0.7,
           color = "black") + 
  scale_y_log10(labels = scales::dollar)+
   scale_fill_manual(values = c("imp" = "orange","exp" = "darkgreen")) +
    theme(text = element_text('xkcd'),
          legend.position = "top")
g + h

```

------------------------------------------------------------------------

Note:

-   The y axis is in log scale, so instead of a <br>1, 2, 3... progression expect a <br>1, 10, 100... progression
-   The data from the point where the export value substantially increases is missing, which is data of 2013.

### Number of trades with high value

```{r}
thematic::thematic_off()

xrange <- c(0,20) # might need to change range
yrange <- c(0,30) # might need to change range
ratioxy <- diff(xrange)/diff(yrange)
mapping <- aes(x, y,
scale,
ratioxy,
angleofspine ,
anglerighthumerus,
anglelefthumerus,
anglerightradius,
angleleftradius,
anglerightleg,
angleleftleg,
angleofneck)

dataman <- data.frame(x= 0, y= 2,
scale = 2,
ratioxy = ratioxy,
angleofspine = -pi/2 ,
anglerighthumerus =  -pi/12,
anglelefthumerus = pi + pi/6,
anglerightradius = pi/2 - pi/4,
angleleftradius = -pi/3,
anglerightleg = 3*pi/2 - pi / 12,
angleleftleg = 3*pi/2 + pi / 12 ,
angleofneck = runif(1, 3*pi/2-pi/10, 3*pi/2+pi/10))

j <- ggplot() + 
  # xkcdaxis(xrange,yrange) +
  xkcdman(mapping,linetype = 3, dataman) + 
  theme_void() +
  # this will allow the text outside of the plot panel
  coord_cartesian(clip = 'off',
                  expand = FALSE) +
  annotate("richtext", x = 3, y = 5.4, ymax = 10,xmin = -4,xmax = 10,  label = "From this figure <br>we see most of the total <br>import trades were of low <br>value, while the <br>export trade trade has<br> almost always been valuable", family = "xkcd",
           fill = "lightgrey") +
  theme(line = element_blank(),
        text = element_text('xkcd')) +
  # annotate("rect", xmin = 2, xmax = 5, ymin = 4.5, ymax = 10,
  # alpha = .2) +
  annotate("segment",x = 0.9, xend = 1.5 ,
             y = 2.5, yend = 3.2) 

i <- clean_coal |> 
    filter(value_usd > 1 ) |> 
    ggplot(aes(value_usd, fill= trade, color = trade)) +
    geom_density(aes(y = ..count..),
                 alpha = 0.7, colour = "black") +
  # geom_histogram(position = "identity",
  #                alpha = 0.5,
  #                binwidth = 1) +
    scale_x_log10(labels = scales::dollar) +
    scale_fill_manual(values = c("imp" = "orange","exp" = "darkgreen")) +
    theme(text = element_text('xkcd'),
          legend.position = "top")

j + i
```

***

Note:

- The trades of value $0 signify represent years where there was no trade value recorded


### Import vs Export over the years

```{r,fig.height=9, fig.width=19}
thematic::thematic_off()
xrange <- c(0,20) # might need to change range
yrange <- c(0,30) # might need to change range
ratioxy <- diff(xrange)/diff(yrange)
mapping <- aes(x, y,
scale,
ratioxy,
angleofspine ,
anglerighthumerus,
anglelefthumerus,
anglerightradius,
angleleftradius,
anglerightleg,
angleleftleg,
angleofneck)

dataman <- data.frame(x= 0, y= 2,
scale = 2,
ratioxy = ratioxy,
angleofspine = -pi/2 ,
anglerighthumerus =  -pi/6 - 135*pi/180,
anglelefthumerus = pi + pi/6,
anglerightradius = -pi/3 - 3*pi/4,
angleleftradius = -pi/3 - pi/2,
anglerightleg = 3*pi/2 - pi / 12,
angleleftleg = 3*pi/2 + pi / 12 ,
angleofneck = runif(1, 3*pi/2-pi/10, 3*pi/2+pi/10))


k <- ggplot() + 
  # xkcdaxis(xrange,yrange) +
  xkcdman(mapping,linetype = 3, dataman) + 
  theme_void() +
  # this will allow the text outside of the plot panel
  coord_cartesian(clip = 'off',
                  expand = FALSE) +
  annotate("richtext", x = 3, y = 5, ymax = 10,xmin = -4,xmax = 10,  label = "Here we can <br>see the progression <br>of the trade values over <br>the years since 2005", family = "xkcd",
           fill = "lightgrey") +
  theme(line = element_blank(),
        text = element_text('xkcd')) +
  # annotate("rect", xmin = 2, xmax = 5, ymin = 4.5, ymax = 10,
  # alpha = .2) +
  annotate("segment",x = 0.9, xend = 1.5 ,
             y = 2.5, yend = 3.2) 


l <- clean_coal |> 
  ggplot(aes(value_usd, year, fill = trade, group = year)) +
  geom_density_ridges(scales = 1,
                      alpha = 0.7) +
  scale_x_log10(labels = scales::dollar) +
    scale_fill_manual(values = c("imp" = "orange","exp" = "darkgreen")) +
    theme(text = element_text('xkcd'),
          legend.position = "top") +
  facet_wrap( ~ trade) # TODO to animate graphs above 



dataman2 <- data.frame(x= 0, y= 2,
scale = 2,
ratioxy = ratioxy,
angleofspine = -pi/2 ,
anglerighthumerus =  -pi/6,
anglelefthumerus =  pi/16,
anglerightradius = pi/16,
angleleftradius = -pi/12 + pi/4,
anglerightleg = 3*pi/2 - pi / 12,
angleleftleg = 3*pi/2 + pi / 12 ,
angleofneck = runif(1, 3*pi/2-pi/10, 3*pi/2+pi/10))


m <- ggplot() + 
  # xkcdaxis(xrange,yrange) +
  xkcdman(mapping,linetype = 3, dataman2) + 
  theme_void() +
  # this will allow the text outside of the plot panel
  coord_cartesian(clip = 'off',
                  expand = FALSE) +
  annotate("richtext", x = 3, y = 5, ymax = 10,xmin = -4,xmax = 10,  label = "Here we can <br>see the progression <br>of the trade values over <br>the years since 2005", family = "xkcd",
           fill = "lightgrey") +
  theme(line = element_blank(),
        text = element_text('xkcd')) +
  # annotate("rect", xmin = 2, xmax = 5, ymin = 4.5, ymax = 10,
  # alpha = .2) +
  annotate("segment",x = 0.9, xend = 1.5 ,
             y = 2.5, yend = 3.2) 
k + l + m
```

------------------------------------------------------------------------

Note:

-   The imports values were way higher in the earlier years as compared to recently, which i think is a good sign
-   The exports values are moving in the opposite direction, which is also a good thing

### The cost of producing coal has gone up over the years .. and so has the value

```{r}
n <- clean_coal |> 
  ggplot(aes(year, coal_rent_pct_of_gdp)) + 
  geom_line() +
  scale_y_continuous(labels = scales::percent)

o <- clean_coal |> 
  ggplot(aes(year, natural_capital)) + 
  geom_line()+
  scale_y_continuous(labels = scales::dollar)

p <- clean_coal |> 
  ggplot(aes(coal_rent_pct_of_gdp, natural_capital, group = as.factor(year))) + 
  geom_point(alpha = 0.7) +
  geom_smooth(group = 1,
              method = "lm",
              color = "darkgrey",
              lty = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_continuous(labels = scales::percent)

dataman2 <- data.frame(x= 0, y= 2,
scale = 2,
ratioxy = ratioxy,
angleofspine = -pi/2 ,
anglerighthumerus =  -pi/6,
anglelefthumerus =  pi/16,
anglerightradius = pi/16,
angleleftradius = -pi/12 + pi/4,
anglerightleg = 3*pi/2 - pi / 12,
angleleftleg = 3*pi/2 + pi / 12 ,
angleofneck = runif(1, 3*pi/2-pi/10, 3*pi/2+pi/10))


q <- ggplot() + 
  # xkcdaxis(xrange,yrange) +
  xkcdman(mapping,linetype = 3, dataman2) + 
  theme_void() +
  # this will allow the text outside of the plot panel
  coord_cartesian(clip = 'off',
                  expand = FALSE) +
  annotate("richtext", x = 3, y = 5, ymax = 10,xmin = -4,xmax = 10,  label = "These two variables <br>showed a high correlation<br> throughout the years", family = "xkcd",
           fill = "lightgrey") +
  theme(line = element_blank(),
        text = element_text('xkcd')) +
  # annotate("rect", xmin = 2, xmax = 5, ymin = 4.5, ymax = 10,
  # alpha = .2) +
  annotate("segment",x = 0.9, xend = 1.5 ,
             y = 2.5, yend = 3.2) 

(n + o) / (q + p)

```



